Wall-to-wall mapping of peat depth from Lidar terrain and airborne radiometrics in Norwegian landscapes

### 2024-10-15 ms/ms.tex
ms: variable importance methods

The last terrain-based predictor we included was the depth-to-water index \citep{murphyMappingWetlandsComparison2007}.
This index approximates a location's vertical height above the surface water feature that it is likely to drain towards.
It is calculated as the minimum cumulative slope (scaled by cell size) to a surface water feature \citep[eq. 5 in][]{murphyTopographicModellingSoil2009}.
We calculated unitless slope from the \unit{1\,m} DTM using the Whitebox software.
Also using Whitebox, we defined surface water features from the DTM by filling depressions and then calculating flow accumulation to define catchment areas for each cell \citep{schonauerSpatiotemporalPredictionSoil2021, schonauerRcodeCalculatingDepthtowater2021}.
This catchments area layer was then thresholded at seven different levels (\emph{flow initiation areas} from \unit{0.25\,ha} to \unit{16\,ha}) to estimate surface water features under moisture scenarios varying from wet to dry \citep{murphyModellingMappingTopographic2011, agrenEvaluatingDigitalTerrain2014, schonauerSpatiotemporalPredictionSoil2021}.
In addition, all surface water features mapped in the AR5 dataset were also transferred to the raster layer.
For each of the seven surface water layers, we derived the depth-to-water index using the \emph{Distance Accumulation} tool in ArcGIS Pro (v.3.1, ESRI, USA), which has an efficient algorithm to find the cumulative distance over a cost surface to the least-cost source.

@@

>>>>

The last terrain-based predictor we included was the depth-to-water index \citep{murphyMappingWetlandsComparison2007}.
This index approximates a location's vertical height above the surface water feature that it is likely to drain towards.
It is calculated as the minimum cumulative slope (scaled by cell size) to a surface water feature \citep[eq. 5 in][]{murphyTopographicModellingSoil2009}.
We calculated unitless slope from the \unit{1\,m} DTM using the Whitebox software.
Also using Whitebox, we defined surface water features from the DTM by filling depressions and then calculating flow accumulation to define catchment areas for each cell \citep{schonauerSpatiotemporalPredictionSoil2021, schonauerRcodeCalculatingDepthwater2021}.
This catchments area layer was then thresholded at seven different levels (\emph{flow initiation areas} from \unit{0.25\,ha} to \unit{16\,ha}) to estimate surface water features under moisture scenarios varying from wet to dry \citep{murphyModellingMappingTopographic2011, agrenEvaluatingDigitalTerrain2014, schonauerSpatiotemporalPredictionSoil2021}.
In addition, all surface water features mapped in the AR5 dataset were also transferred to the raster layer.
For each of the seven surface water layers, we derived the depth-to-water index using the \emph{Distance Accumulation} tool in ArcGIS Pro (v.3.1, ESRI, USA), which has an efficient algorithm to find the cumulative distance over a cost surface to the least-cost source.

@@

We quantified global variable importance for the RFs with the 25 quantitative predictors.
Global variable importance measures the influence of a given predictor on the output of the model, aggregated across all locations.
We calculated variable importance with the \emph{vip} R package (v.0.4.1), by three different methods: \emph{FIRM}, \emph{permutation}, and \emph{Shapley} \citep{greenwellVariableImportancePlots2020}.
\emph{Permutation} values were aggregated from ten iterations with RMSE as the performance measure.
